Network based multifactorial modelling microRNA-target interaction.

### 2019-09-17 article.Rmd
revision of SERPINE changes

We have applied our model in a real dataset for which experimental measurements of various factors were available. Expression levels of miRNA and genes in tumor and normal tissue of single patient are retrieved from TCGA Research Network: https://www.cancer.gov/tcga. High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015]. We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from published datasets [@helwak_mapping_2013; @moore_mirnatarget_2015]. Detailed description of network construction and its code is available in Supplementary Materials and Methods. We have determined the most effective nodes in network and fold changes of genes as compared tumor and normal tissue samples of the same patient.As a result of comparison of these two datasets, we selected a gene , SERPINE2, as trigger for simulation of model (see Supplementary Materials and Methods).

@@

[#TODO]: <> "figure1, 2 ve 3 uzerinden farkli dagilim ve degradasyon açıklaması buraya. G2 uzerinden olmasa iyi olur" 
[FIXED by selcen]
When the model run based on expression values of miRNAs and ceRNAs (genes), proportional distribution of miRNA on targets is determined in accordance with \eqref{eq:1, eq:2}. In that case, the system is observed as shown in \autoref{fig1}a at steady-state. However, if interaction factors are taken into account in the approach, specified interaction factors (i.e, binding factors such as seed type and energy) affect distribution of miRNA on targets as calculated in \eqref{eq:3} (shown in \autoref{fig2}c). After distributed miRNA efficiency was re-calculated with degradation factor(s) (i.e region effect) according to \eqref{eq:4}, steady-state condition of network is obtained (shown in \autoref{fig2}d and \autoref{fig3}a). For instance, proportional distribution of G1:M1 interaction in \autoref{fig1}a is calculated lower than the same pair (G1:M1) in the other approach that includes interaction factors (\autoref{fig2}c). But, a decrease is observed in repression efficiency of M1 miRNA on G1 in interaction factor based network approach (\autoref{fig2}d and \autoref{fig3}a).

@@

[#TODO] We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach. Simulation results were not expected to be consistent with tumor tissue expression values because large number of abnormal regulations were observed in these tissues. So, it might not said that a single gene is responsible for entire regulations in cancer cells. Besides, the all other factors such as up/down regulation of miRNAs or other genes were ignored. Therefore, it would be more useful to test our approach with an ideal dataset that includes entire informations of gene expression values before and after the regulation of a known gene at existing tissue, and miRNA expression values of this tissue at initial conditions.

[#TODO]: <> "run perturbation screening on tcga set, find a gene then couple sentences about that gene"
[fixed by selcen]

>>>>

We have applied our model in a real dataset for which experimental measurements of various factors were available. Expression levels of miRNA and genes in tumor and normal tissue of single patient are retrieved from TCGA Research Network: https://www.cancer.gov/tcga. High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015]. We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from published datasets [@helwak_mapping_2013; @moore_mirnatarget_2015]. Detailed description of network construction and its code is available in Supplementary Materials and Methods. SERPINE', one of the most effective nodes, has been used as trigger in ceRNA network simulation in breast cancer patient network. Relevant steps and codes are available at  Supplementary Materials and Methods.

@@

[#TODO]: <> "figure1, 2 ve 3 uzerinden farkli dagilim ve degradasyon açıklaması buraya. G2 uzerinden olmasa iyi olur"

When the model run based on expression values of miRNAs and ceRNAs (genes), proportional distribution of miRNA on targets is determined in accordance with \eqref{eq:1, eq:2}. In that case, the system is observed as shown in \autoref{fig1}a at steady-state. However, if interaction factors are taken into account in the approach, specified interaction factors (i.e, binding factors such as seed type and energy) affect distribution of miRNA on targets as calculated in \eqref{eq:3} (shown in \autoref{fig2}c). After distributed miRNA efficiency was re-calculated with degradation factor(s) (i.e region effect) according to \eqref{eq:4}, steady-state condition of network is obtained (shown in \autoref{fig2}d and \autoref{fig3}a). For instance, proportional distribution of G1:M1 interaction in \autoref{fig1}a is calculated lower than the same pair (G1:M1) in the other approach that includes interaction factors (\autoref{fig2}c). But, a decrease is observed in repression efficiency of M1 miRNA on G1 in interaction factor based network approach (\autoref{fig2}d and \autoref{fig3}a).

@@

We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach. Simulation results were not expected to be consistent with tumor tissue expression values because large number of abnormal regulations were observed in these tissues. So, it might not said that a single gene is responsible for entire regulations in cancer cells. Besides, the all other factors such as up/down regulation of miRNAs or other genes were ignored. Therefore, it would be more useful to test our approach with an ideal dataset that includes entire informations of gene expression values before and after the regulation of a known gene at existing tissue, and miRNA expression values of this tissue at initial conditions.
